---
title: "Task 3"
author: "Amelia Ritger"
date: "2/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

#load packages
library(tidyverse)
library(tidyverse)
library(here)
library(janitor)
library(raster)
library(sf)
library(tmap)
#library(tmaptools)
library(gstat)

library(plotly)
library(paletteer)
```

My task?

Make finalized map(s) in which I: 

- Clearly and professionally show the different watersheds for all main Hawaiian islands, or for an island or region of my choosing

Do you want your maps to be static or interactive?
What areas do you want to focus on? 
Do you want to show land use/landcover and watersheds in a single map or in separate maps? 
Just make sure that your final outputs are presented in a nice professional HTML that you’d be proud to share with someone so that they can see your awesome spatial code! 

Let's get cracking!

Read in data
```{r}
water <- read_sf(here("Watersheds", "Watersheds.shp"))
```

Basically plot it
```{r}
plot(water)
```

Unfortunately, swma (water management area) doesn't look very interesting...
"wuc" (watershed unit code text) and "huc" (watershed unit code numeric) also don't look too interesting

Check on other map related things and map it!
```{r}
st_crs(water) #cool, EPSG is 4326, as it should be

#Rename variable of interest in the name of good-looking maps
water <- water %>% 
  mutate(`Watershed Area (sq m)`= hucarea)

water_map <- ggplot(data=water) +
  geom_sf(aes(fill=`Watershed Area (sq m)`, color=`Watershed Area (sq m)`), show.legend = FALSE) +
  scale_fill_paletteer_c("scico::hawaii", direction=-1) +
  scale_color_paletteer_c("scico::hawaii", direction=-1) +
  ggtitle("Map of the watersheds of the Hawaiian Islands") +
  xlab("Longitude") +
  ylab("Latitude") +
  theme(panel.grid.major = element_line(color = "gray"(0.5), linetype = "dashed", size = 0.5),
        panel.background = element_rect(fill = "aliceblue"))

ggplotly(water_map) %>%
  highlight(
    "plotly_hover",
    selected = attrs_selected(line = list(color = "black")))
```

